Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-04-19 apj.tex
more modifications for Greg's comments

\color{orange}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in a linear model these values are arbitrary.
\color{orange}
As a technical aside whose effect will be noted later, we highlight a choice of model parameterization that enables the efficient evaluation of the
posterior by the software package
STAN \citep{stan}.  Given $N$ supernovae, each of the above conditions reduce the number of degrees of freedom by one,
for example  the $N$ $\Delta$ values are specified by $N-1$ independent parameters. 
Our implementations
with $N-1$ independent parameters do not converge efficiently, presumably because the highly correlated parameters result in
no single ``mass'' in the Hamiltonian Monte Carlo being appropriate over the full parameter space.  
Based on the recommendation in the STAN User's Guide
and reinforced by our own experience, each of the above vectors is described by $N$ parameters.  While this introduces an unconstrained
degree of freedom, it improves the efficiency in evaluating the posterior in the other parameters of interest.

We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

@@

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
\color{orange}
Note that the extreme outlier has large uncertainties, and is a manifestation of the unconstrained degree of freedom in this parameter set.
\color{black}
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$
%\color{orange}
%While at low  $\gamma^0_V k_0$
%\color{black}
%On the other hand, lower values of $\gamma^1_V k_1$
%occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
%\citet{2011ApJ...729...55F} and the above authors
%find that supernovae with large extinction 
%preferentially exhibit low $R^F$.

@@

\subsection{Host-Galaxy Mass}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that when introducing intrinsic color scatter as a latent parameter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore whether the Hubble residual may be due to unaccounted color,
we plot in Figure~\ref{childress:fig} our intrinsic parameter  $A_{\delta I}$  versus host mass 
\citep{2016rigault}, and their ideograms for two samples divided at host mass of  $\log{(M/M_\sun)}=10$
as in  \citet{2013ApJ...770..108C}.
Although by eye it appears that the high-mass hosts tend the have lower $A_{\delta I}$, the difference is
not significant; we calculate  average
%---
$\langle A_{\delta I} \rangle=    -0.0026 \pm    0.0010$,
$\langle A_{\delta I} \rangle=  -0.0037 \pm    0.0010 $ 
for low- and high-mass hosts respectively. 
Using the SNfactory sample,
\citet{2013A&A...560A..66R,2017Rigault} find that a step in Hubble residuals is better related to local star formation rate, rather than
global host mass.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/rigault3.pdf}
s     \includegraphics[width=2.8in]{output23/rigault2.pdf}
      \caption{Left: Intrinsic parameter $A_{\delta I}=\delta_I D$  versus host galaxy mass. Overplotted are the mean and 1$\sigma$ values for supernovae with hosts
      less than and greater than  $\log{(M/M_\sun)}=10$.
Right: Ideograms of  $A_{\delta I}$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
   \label{childress:fig}}
\end{figure}

\subsection{Physical Motivation}

>>>>

\color{orange}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in a linear model these values are arbitrary.
\color{orange}
We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

@@

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$
%\color{orange}
%While at low  $\gamma^0_V k_0$
%\color{black}
%On the other hand, lower values of $\gamma^1_V k_1$
%occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
%\citet{2011ApJ...729...55F} and the above authors
%find that supernovae with large extinction 
%preferentially exhibit low $R^F$.

@@

\subsection{Host-Galaxy Mass}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that when introducing intrinsic color scatter as a latent parameter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore whether the Hubble residual may be due to unaccounted color,
we plot in Figure~\ref{childress:fig} our intrinsic parameter  $A_{\delta I}$  versus host mass 
\color{orange}
for the subset of supernovae whose hosts are given in \citet{2013ApJ...770..108C}.
Also plotted  are ideograms of $A_{\delta I}$ for the two subsamples obtained by dividing the full sample at a host mass of  $\log{(M/M_\sun)}=10$.
Although by eye it appears that the high-mass hosts tend the have lower $A_{\delta I}$, the difference is
not significant; we calculate  average
%---
$\langle A_{\delta I} \rangle=   -0.0026 \pm    0.0014$,
$\langle A_{\delta I} \rangle=-0.0035 \pm    0.0010$ 
for low- and high-mass hosts respectively. 
Using the SNfactory sample,
\citet{2013A&A...560A..66R,2017Rigault} find that a step in Hubble residuals is better related to local star formation rate, rather than
global host mass.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/childress.pdf}
s     \includegraphics[width=2.8in]{output23/childress2.pdf}
      \caption{Left: Intrinsic parameter $A_{\delta I}=\delta_I D$  versus host galaxy mass. Overplotted are the mean and 1$\sigma$ values for supernovae with hosts
      less than and greater than  $\log{(M/M_\sun)}=10$.
Right: Normalized ideograms of  $A_{\delta I}$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
   \label{childress:fig}}
\end{figure}

\color{black}
\subsection{Physical Motivation}
